---
title: "Aids Drug"
author: "Danitza Nunez"
date: "28 September 2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
Drugees=filter(triUmp,DrugUser=="yes")
Females=filter(triUmp,gender=="F")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aids drug (triUmpsal) is under test in clinical trials. Consider the DrugUsers, does this drug seem effective on this group?  Hint - use dplyr to obtain that sub-population with the filter command.  Next, repeat the analysis on just females, does this drug seem effective on women?


<!-- Be sure to produce Descriptive Results and Inferential Results for both groups. !-->



1. Is the drug effective for Drug Users?

$H_0$: There is no difference in the drug  effectiveness in drug users.  
$H_A$: There is a significant difference of effectiveness in drug users. 


2. Is the vaccine effective for women?

$H_0$ : There is no effectiveness in the vaccine for females.  
$H_A$: There is a significant difference in effectiveness in the vaccine for Females.  


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed, for example sub-setting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

The helper file being used to determine the effectiveness of drug use in drug users and vaccines for Females is CAT~ CAT

# Results

<!--Divide this section into two sub-sections:  One for your descriptive results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe patterns, if any, that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming to or departing from what is expected by the null hypothesis !-->



### Drug users
```{r}
barchartGC(~infected + treatment,data=Drugees)
barchartGC(~infected + treatment,data=Drugees, type="percent")

```


The graphical descriptive results show that  Drug users treated with the vaccine are 4 times less likely to get AIDS. 


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns, if any, that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=Drugees)
rowPerc(table1)
colPerc(table1)
```


Among all drug users who were infected with AIDS, 24% of them had received the vaccine and about 76% of those who were infected received the placebo. This suggests drug users who received the vaccine were less likely to get AIDS. 




## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfies the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
The p-value is 0.0000001242 which is smaller than 0.05, so we reject the null hypothesis. This shows the vaccine is effective in the drug user community.  

Among those who did not received the treatment drug, they are about 0.35 times more likely to contract aids, than those who recieved treatment. 

## Descriptive Results for Females


### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe patterns, if any, that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming to or departing from what is expected by the null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

Females treated with the vaccine are more susceptible to get AIDS. 


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns, if any, that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~treatment + infected, data=Females)
rowPerc(table1)
colPerc(table1)
```


Among all females who were infected with AIDS, 58% of them had received the vaccine and 48% of those who were infected received the placebo. This suggests females who received the vaccine were more susceptible to get AIDS.   



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfies the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


In the Chi-Squared Test The p-value is 0.00007686 which is smaller than 0.05, so we reject the null hypothesis. In the Fisher test the p-value is p-value = 6.596e-05. This shows the vaccine is not effective in the female community.

if you are a woman you are 1.54 times more likely to contract AIDS if you took placebo than those if you got the vaccine. 


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the George Town videos collection.!

Report on both Drug Users and Females.  Be sure to include the odds for both groups.
-->


1. Based on the following, it demonstrates drug users who have received the treatment vaccine are less susceptible to get infected with AIDS. The vaccine is effective for drug users. We reject the null hypothesis.

2. Based on the following, it demonstrates females who have received the treatment vaccine are more susceptible to get infected with AIDS. The vaccine is not effective for females. We reject the null hypothesis. 
 
